---
title: "Marijuana Legalization Simulation" 
description: |
  Analysis of Proportion of People who Support/Oppose the Legalization of Marjiuana
author: Julia Aceron
date: November 8, 2024
format: html
execute:
  warning: false
  message: false
---


```{r}
library(tidyverse)
library(openintro)
library(purrr)
```
# Simulation of those who support/oppose 

## Code For Simulation

#### This dataset shows the number of people who chose to oppose or support the legalization of marijuana. This simulation aims to determine if there is a difference between the proportion of supporters/opposers. The null hypothesis is that there is no difference between the proportion of people who support the legalization of marijuana and the proportion of people who oppose the legalization (50% of people support either side). The alternative hypothesis is that there is a difference between the two groups. 

```{r}
set.seed(47) 

support_obs <- leg_mari |> 
  summarize(prop_support = mean(response == "support")) |>
  pull()

random_response <- function(rep, num_people) {
  response = sample(c("support", "oppose"), size = num_people, replace = TRUE, prob = c(0.5, 0.5))
  return(mean(response == "support"))
}

num_sim <- 4747
support_random <- map_dbl(1:num_sim, random_response, num_people = 47)

support_random |>
  data.frame() |> 
  ggplot(aes(x = support_random)) +
  geom_histogram() +
  labs(
    title = "Sampling Distribution When Null is True",
    subtitle = "no inherent preference for those who oppose or support marijuana",
    x = "Proportion of People who Support Marijuana"
  )

p_value <- sum(support_random >= support_obs)/num_sim
  
statistics_table <- data.frame(p_value, support_obs)
statistics_table
```

## Process 

#### First, the observed statistic for the number of people who supported the proposition as about 0.697. Then, a simulation was conducted in which that the null hypothesis was assumed. In other words, the proportions were 0.5 for supporters and opposers to see if getting a statistic of 0.697 was by chance alone. After running this simulation many, many times and graphing the results, the sampling distribution showed the probabilties of getting different observed statistics. 

## Analysis 

#### From the graph, we can see that getting the observed statistic of about 0.697 results in a p-value of 0.004. Since the p-value is less than the standarized value of 0.05, we can reject the null hypothesis stating that 50% of the 119 survey respondents between the ages of 18 and 34 would support or oppose the legalization of marijuana. We can therefore conclude that there is evidence that there is a difference between these two proportions since getting a statistic of 0.697 for those who support marijuana occurs rarely with luck. 

Source: 
This dataset was from the USA Survey, collected from July 8-11. 
[Legalization of Marijuana Support in 2010 California Survey](https://github.com/OpenIntroStat/openintro/blob/HEAD/R/data-leg_mari.R) 

