---
title: "Analysis of Broadway Show Synopses and their Effect on Weekly Grosses"
author: "Julia Aceron"
subtitle: "December 6th, 2024"
format:
  revealjs:
    scrollable: true
    slide-number: true
    show-slide-number: all
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

```{r, echo = FALSE}
library(tidyverse)
library(stringr)
```


```{r, echo = FALSE}
grosses <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-28/grosses.csv', guess_max = 40000)
synopses <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-28/synopses.csv')
cpi <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-28/cpi.csv')
pre_1985_starts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-28/pre-1985-starts.csv')
```

```{r, echo = FALSE}
shows_combined <- 
  dplyr::full_join(grosses, synopses, by = "show")
```

## Road Map 
- Effect of 'Musical' keyword 
- Effect of 'Play' keyword
- Effect of Shows Featuring Artistic Characters
- Effect of Synopsis Lengths 

## Why this matters 
- Important to know what sells 
- What playwrights can focus on in the future 

## Effect of Words in a Synopsis on Weekly Gross 


```{r, echo = FALSE}

synopsis_musical_count <- shows_combined |>
  mutate(musical_count = str_count(synopsis, "(?i)\\bmusical\\b")) |> # Counting the number of times 'musical' appears in the synopsis
  select(show, weekly_gross, synopsis, musical_count) # Selecting these specific columns in the dataframe

synopsis_musical_count |> 
  filter(!is.na(musical_count)) |> # Filtering for values that are not NA
  mutate(weekly_gross = weekly_gross/1000) |> # Creating a new column that puts the gross in thousands 
  select(musical_count, weekly_gross) |> # Selecting these columns
  arrange(desc(musical_count)) |> # Arranging in descending order by musical_count
  ggplot(aes(x = as.factor(musical_count), y = weekly_gross)) + # Defining the variables and ordering musical_count 
  geom_boxplot(fill = "pink", color = "black", outlier.color = "lightgreen") + # Assigning colors
  labs (
    title = "Relationship Between 'Musical' Count in Synopsis and Weekly Gross",
    subtitle = "Influence of 'Musical' keyword frequency", 
    x = "Number of Times 'Musical' Appears in Synopsis", 
    y = "Weekly Gross Revenue (in thousands)"
  ) + # Renaming variables for clarity
  theme_minimal(base_size = 14) + # Theme 
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),  # Centering and bold the title
    plot.subtitle = element_text(hjust = 0.5, size = 12, face = "italic"),  # Subtitle formatting
    axis.title = element_text(face = "bold"),  # Bolding axis labels
    axis.text = element_text(size = 12),  # Making axis text bigger
    panel.grid.major = element_line(color = "gray", size = 0.5),  # Lighter gridlines
    panel.grid.minor = element_blank(), # No minor gridlines
  )
```
## Effect of Words in a Synopsis on Weekly Gross 

```{r, echo = FALSE}
synopsis_play_count <- shows_combined |>
  mutate(play_count = str_count(synopsis, "(?i)\\bplay\\b")) |>
  select(show, weekly_gross, synopsis, play_count)

synopsis_play_count |>
  filter(!is.na(play_count)) |>
  mutate(weekly_gross = weekly_gross/1000) |>
  select(play_count, weekly_gross) |>
  arrange(desc(play_count)) |>
  ggplot(aes(x = as.factor(play_count), y = weekly_gross)) +
  geom_boxplot(fill = "pink", color = "black", outlier.color = "lightgreen") +
  labs (
    title = "Relationship Between 'Play' Count in Synopsis and Weekly Gross",
    subtitle = "Influence of 'Play' keyword frequency", 
    x = "Number of Times 'Play' Appears in Synopsis", 
    y = "Weekly Gross Revenue (in thousands)"
  ) +
  theme_minimal(base_size = 14) + # Theme 
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),  # Centering and bold the title
    plot.subtitle = element_text(hjust = 0.5, size = 12, face = "italic"),  # Subtitle formatting
    axis.title = element_text(face = "bold"),  # Bolding axis labels
    axis.text = element_text(size = 12),  # Making axis text bigger
    panel.grid.major = element_line(color = "gray", size = 0.5),  # Lighter gridlines
    panel.grid.minor = element_blank(), # No minor gridlines
  )
```

## Effect of Shows Featuring Artistic Characters on Weekly Gross

```{r, echo = FALSE}
artist_characters <- shows_combined |>
  mutate(synopsis_lower = str_to_lower(synopsis),
         is_artist = str_detect(synopsis_lower, "(singer|playwright|actor|chorus|dancer)")) |>
  select(show, weekly_gross, synopsis, is_artist)


artist_characters |>
  filter(!is.na(is_artist)) |>
  mutate(weekly_gross = weekly_gross/1000) |>
  select(is_artist, weekly_gross) |>
  arrange(desc(is_artist)) |>
  ggplot(aes(x = is_artist, y = weekly_gross)) +
  geom_boxplot(fill = "pink", color = "black", outlier.color = "lightgreen") +
  labs (
    title = "Shows with Artistic Characters vs Weekly Gross",
    subtitle = "Impact of Artistic Themes on Revenue",
    x = "Contains Artistic Characters", 
    y = "Weekly Gross Revenue"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),  # Centering and bold the title
    plot.subtitle = element_text(hjust = 0.5, size = 12, face = "italic"),  # Subtitle formatting
    axis.title = element_text(face = "bold"),  # Bolding axis labels
    axis.text = element_text(size = 12),  # Making axis text bigger
    panel.grid.major = element_line(color = "gray", size = 0.5),  # Lighter gridlines
    panel.grid.minor = element_blank(), # No minor gridlines
  )
```

## Effect of Synopsis Lengths on Weekly Gross 

```{r, echo = FALSE}
words_synopsis <- shows_combined |>
  mutate(synopsis_count = str_count(synopsis, "\\S+")) |>
  select(show, weekly_gross, synopsis, synopsis_count)

words_synopsis |>
  mutate(weekly_gross = weekly_gross/1000) |> 
  mutate(synopsis_bin = cut(synopsis_count, breaks = seq(0, max(synopsis_count, na.rm = TRUE), by = 50))) |>
  group_by(synopsis_bin) |>
  filter(!is.na(synopsis_bin)) |>
  head(10000) |>
  ggplot(aes(x = synopsis_bin, y = weekly_gross)) +
  geom_col(fill = "pink") +
  labs (
    title = "Top Synopsis Lengths vs Weekly Gross",
    x = "Synopsis Length", 
    y = "Weekly Gross"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),  # Centering and bold the title
    plot.subtitle = element_text(hjust = 0.5, size = 12, face = "italic"),  # Subtitle formatting
    axis.title = element_text(face = "bold"),  # Bolding axis labels
    axis.text = element_text(size = 12),  # Making axis text bigger
    panel.grid.major = element_line(color = "gray", size = 0.5),  # Lighter gridlines
    panel.grid.minor = element_blank(), # No minor gridlines
  )
```

## Conclusion 
- No significant difference
- Playwrights should just express their creativity

## Thank you!

## Source
TidyTuesday Dataset

[Broadway Weekly Grosses from Playbill. Includes data on revenue and attendance figures for theatres that are a part of The Broadway League.](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-04-28/readme.md) 

