---
title: "Mean Absorbance by Publication Data Analysis" 
description: |
  Analysis of Mean Absorbance and Frequency
author: Julia Aceron
date: November 26, 2024
format: html
execute:
  warning: false
  message: false
---

## PLAN 
#### In this analysis, I plan to use the WAI database and explore the following tables: Measurements, PI_Info, and Subjects. To correctly analyze mean absorbance and frequency with each publication in this database, I would have to join these tables so that the information from the Measurements table is correlated with the publications needed in PI_Info. The Subjects table will be used to create the legend of the plot. 
```{r}
library(tidyverse)
library(dplyr)
library(DBI)
library(shiny)
```

## Create SQL Connection
```{r}
library(RMariaDB)
con_wai <- DBI::dbConnect(
  RMariaDB::MariaDB(), host = "scidb.smith.edu",
  user = "waiuser", password = "smith_waiDB", 
  dbname = "wai"
)
Measurements <- tbl(con_wai, "Measurements")
PI_Info <- tbl(con_wai, "PI_Info")
Subjects <- tbl(con_wai, "Subjects")

# collect(Measurements)
```

## Show Database Tables
```{sql, connection=con_wai}
SHOW TABLES;
```




## Create table for the plot
```{sql, connection=con_wai}
#| output.var: "graph_info" 

SELECT PI_Info.Identifier, PI_Info.AuthorsShortList, Frequency, AVG(Absorbance) AS mean_absorbance, PI_Info.Year
FROM PI_Info
LEFT JOIN Measurements ON PI_Info.Identifier = Measurements.Identifier
WHERE Measurements.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015", "Lewis_2015", "Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013", "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010") AND Frequency < 8000 AND Frequency > 200
GROUP BY Identifier, Frequency;
```
## Create legend for the plot
```{sql, connection=con_wai}
#| output.var: "labels"
SELECT AuthorsShortList, Year, COUNT(DISTINCT SubjectNumber, Ear) AS N, Instrument, Frequency, AVG(Absorbance) AS mean_absorbance, CONCAT(AuthorsShortList, " (",Year,") ", "N=", COUNT(DISTINCT SubjectNumber, Ear), "; ", Instrument) AS legend
FROM PI_Info AS p
LEFT JOIN Measurements AS m ON m.Identifier = p.Identifier 
WHERE p.Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015", "Lewis_2015", "Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013", "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010") AND Frequency < 8000 AND Frequency > 200
GROUP BY p.Identifier, m.Instrument, Frequency

```
## Plot
```{r}
labels |> ggplot(aes(x = Frequency, y = mean_absorbance, color = legend, group = legend)) +
  labs(
    title = "Mean absorbance from each publication in WAI database",
    x = "Frequency (Hz)", 
    y = "Mean Absorbance",
    color = NULL
  ) +
  geom_line() +
  scale_x_log10(
    breaks = c(200, 400, 600, 800, 1000, 2000, 4000, 6000, 8000),
    labels = c("200", "400", "600", "800", "1000", "2000", "4000", "6000", "8000") 
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    legend.text = element_text(size = 5),
    legend.background = element_rect(fill = "white", color = "black"),
    legend.position = c(0.2,0.85),
    legend.box.margin = margin(0, 0, 0, 0),      
    legend.key.size = unit(0.2, "cm"),
    legend.key.height = unit(0.2, "cm"),
    legend.box = "horizontal"
  ) 
```
## DESCRIPTION 
#### To create the graph above, there were two main steps that I took. The first step was to join the tables Measurements and PI_Info by the Identifier column so that I have a new table that describes the frequencies and mean absorbances for each publication. The second step was to create the legend. I joined the same tables again so that the new table would show the author's label, year, subject size, and corresponding instrument for each publication. As a result, I created a plot that demonstrates the mean absorbance for all of the selected publications depending on the frequencies. 

## Count # of Subjects for each group of interest in each publication
```{sql, connection=con_wai}
SELECT Sex, Race, Ethnicity, Identifier, SubjectNumber,
COUNT(*) 
FROM Subjects 
WHERE Identifier IN ("Abur_2014", "Feeney_2017", "Groon_2015", "Lewis_2015", "Liu_2008", "Rosowski_2012", "Shahnaz_2006", "Shaver_2013", "Sun_2016", "Voss_1994", "Voss_2010", "Werner_2010")
GROUP BY Sex, Race, Ethnicity, Identifier, SubjectNumber
```
## Get frequency and mean absorbance for race
```{sql, connection=con_wai}
#| output.var: "plot_race"

SELECT s.Race, Frequency, s.Identifier, AVG(Absorbance) AS mean_absorbance
FROM Subjects AS s
LEFT JOIN Measurements AS m ON m.SubjectNumber = s.SubjectNumber
WHERE s.Identifier IN ("Rosowski_2012")
GROUP BY s.Race, Frequency, s.Identifier;
```

```{sql, connection=con_wai}
#| output.var: "plot_sex"

SELECT s.Sex, Frequency, s.Identifier, AVG(Absorbance) AS mean_absorbance
FROM Subjects AS s
LEFT JOIN Measurements AS m ON m.SubjectNumber = s.SubjectNumber
WHERE s.Identifier IN ("Rosowski_2012")
GROUP BY s.Sex, Frequency, s.Identifier;
```

```{sql, connection=con_wai}
#| output.var: "plot_ethnicity"

SELECT s.Ethnicity, Frequency, s.Identifier, AVG(Absorbance) AS mean_absorbance
FROM Subjects AS s
LEFT JOIN Measurements AS m ON m.SubjectNumber = s.SubjectNumber
WHERE s.Identifier IN ("Rosowski_2012")
GROUP BY s.Ethnicity, Frequency, s.Identifier;
```

## Plots
```{r}
plot_race |> 
  ggplot(aes(x = Frequency, y = mean_absorbance, color = Race)) +
geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(
    title = "Mean Absorbance by Frequency (Faceted by Race)",
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    color = ""
  ) +
  scale_x_log10() +
  theme_minimal() +
  facet_wrap(~ Race)
  
```

```{r}
plot_sex |> 
  ggplot(aes(x = Frequency, y = mean_absorbance, color = Sex)) +
geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(
    title = "Mean Absorbance by Frequency",
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    color = ""
  ) +
  scale_x_log10() +
  theme_minimal() 
```

```{r}
plot_ethnicity |> 
  ggplot(aes(x = Frequency, y = mean_absorbance, color = Ethnicity)) +
geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(
    title = "Mean Absorbance by Frequency (Faceted by Ethnicity)",
    x = "Frequency (Hz)",
    y = "Mean Absorbance",
    color = ""
  ) +
  scale_x_log10() +
  theme_minimal() +
  facet_wrap(~ Ethnicity)

```




```{r}
dbDisconnect(con_wai)
```


Source: 
This dataset was from the WAI Database. 
[Wideband Acoustic Immittance (WAI) Database](https://www.science.smith.edu/wai-database/) 

[Voss, SE. 2019. “Resource Review.” Ear and Hearing 40 (6).](https://journals.lww.com/ear-hearing/citation/2019/11000/resource_review.21.aspx)

